Assignment 
Homework 1 Exercises
---
Q2.1
(a)
```{r}
# 95th percentile
qnorm(0.95, mean=10, sd=3)
# 99th percentile
qnorm(0.99, mean=10, sd=3)
```
(b)
```{r}
# 95th percentile of the t-distribution with 10 degrees of freedom
qt(0.95, df=10)
# 99th percentile of the t-distribution with 10 degrees of freedom
qt(0.99, df=10)
# 95th percentile of the t-distribution with 25 degrees of freedom
qt(0.95, df=25)
# 99th percentile of the t-distribution with 25 degrees of freedom
qt(0.99, df=25)
```
(c)
```{r}
# 95th percentile of the chi-squared distribution with 1 degrees of freedom
qchisq(0.95, df=1)
# 99th percentile of the chi-squared distribution with 1 degrees of freedom
qchisq(0.99, df=1)

# 95th percentile of the chi-squared distribution with 4 degrees of freedom
qchisq(0.95, df=4)
# 99th percentile of the chi-squared distribution with 4 degrees of freedom
qchisq(0.99, df=4)

# 95th percentile of the chi-squared distribution with 10 degrees of freedom
qchisq(0.95, df=10)
# 99th percentile of the chi-squared distribution with 10 degrees of freedom
qchisq(0.99, df=10)
```
(d)
```{r}
# 95th percentile of the F-distribution with 2 and 10 degrees of freedom
qf(0.95, df1=2, df2=10)
# 99th percentile of the F-distribution with 2 and 10 degrees of freedom
qf(0.99, df1=2, df2=10)

# 95th percentile of the F-distribution with 4 and 10 degrees of freedom
qf(0.95, df1=4, df2=10)
# 99th percentile of the F-distribution with 4 and 10 degrees of freedom
qf(0.99, df1=4, df2=10)
```
---
Q2.2
(a)
```{r}
# 95th percentile of square of standard normal distribution
qnorm(0.95, mean = 0, sd=1)^2

# 90th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.90, df=1)

# 97.5th percentile of square of standard normal distribution
qnorm(0.975, mean = 0, sd=1)^2

# 95th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.95, df=1)

# 98.75th percentile of square of standard normal distribution
qnorm(0.9875, mean = 0, sd=1)^2

# 97.5th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.975, df=1)

```
We can see from the following results that:

1. the 90th percentile of the chi-squared distribution is equal to the 95th percentile of the square of the standard normal distribution.

2. the 95th percentile of the chi-squared distribution is equal to the 97.5th percentile of the square of the standard normal distribution.

3. the 97.5th percentile of the chi-squared distribution is equal to the 98.75th percentile of the square of the standard normal distribution.

(b)
```{r}
v1 <- 1
v2 <- 20

# 95th percentile of square of t distribution
qt(0.95, df=v2)^2

# 90th percentile of f distribution (1, v2)
qf(0.90, df1=v1, df2=v2)

# 97.5th percentile of square of t distribution
qt(0.975, df=v2)^2

# 95th percentile of f distribution (1, v2)
qf(0.95, df1=v1, df2=v2)

# 98.75th percentile of square of t distribution
qt(0.9875, df=v2)^2

# 97.5th percentile of f distribution (1, v2)
qf(0.975, df1=v1, df2=v2)
```
We can see from the following results that:

1. the 90th percentile of the F distribution (1, v2) is equal to the 95th percentile of the square of the t distribution (v2).

2. the 95th percentile of the F distribution (1, v2) is equal to the 97.5th percentile of the square of the t distribution (v2).

3. the 97.5th percentile of the F distribution (1, v2) is equal to the 98.75th percentile of the square of the t distribution (v2).
```{r}
qnorm(0.95, mean = 0, sd=1)^2
qnorm(0.975, mean = 0, sd=1)
```
