---
Title: STA3001 Homework 1
Author: Zane Yee Sun
ID: 12440004
output: pdf_document
---
```{r setup, include = FALSE}
```
Q2.1
(a)
```{r}
# 95th percentile
qnorm(0.95, mean=10, sd=3)
# 99th percentile
qnorm(0.99, mean=10, sd=3)
```
Q2.1
(b)
```{r}
# 95th percentile of the t-distribution with 10 degrees of freedom
qt(0.95, df=10)
# 99th percentile of the t-distribution with 10 degrees of freedom
qt(0.99, df=10)
# 95th percentile of the t-distribution with 25 degrees of freedom
qt(0.95, df=25)
# 99th percentile of the t-distribution with 25 degrees of freedom
qt(0.99, df=25)
```
Q2.1
(c)
```{r}
# 95th percentile of the chi-squared distribution with 1 degrees of freedom
qchisq(0.95, df=1)
# 99th percentile of the chi-squared distribution with 1 degrees of freedom
qchisq(0.99, df=1)

# 95th percentile of the chi-squared distribution with 4 degrees of freedom
qchisq(0.95, df=4)
# 99th percentile of the chi-squared distribution with 4 degrees of freedom
qchisq(0.99, df=4)

# 95th percentile of the chi-squared distribution with 10 degrees of freedom
qchisq(0.95, df=10)
# 99th percentile of the chi-squared distribution with 10 degrees of freedom
qchisq(0.99, df=10)
```
Q2.1
(d)
```{r}
# 95th percentile of the F-distribution with 2 and 10 degrees of freedom
qf(0.95, df1=2, df2=10)
# 99th percentile of the F-distribution with 2 and 10 degrees of freedom
qf(0.99, df1=2, df2=10)

# 95th percentile of the F-distribution with 4 and 10 degrees of freedom
qf(0.95, df1=4, df2=10)
# 99th percentile of the F-distribution with 4 and 10 degrees of freedom
qf(0.99, df1=4, df2=10)
```
Q2.2
(a)
```{r}
# 95th percentile of square of standard normal distribution
qnorm(0.95, mean = 0, sd=1)^2

# 90th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.90, df=1)

# 97.5th percentile of square of standard normal distribution
qnorm(0.975, mean = 0, sd=1)^2

# 95th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.95, df=1)

# 98.75th percentile of square of standard normal distribution
qnorm(0.9875, mean = 0, sd=1)^2

# 97.5th percentile of chi-squared distribution with 1 degree of freedom
qchisq(0.975, df=1)

```
We can see from the following results that:

1. the 90th percentile of the chi-squared distribution is equal to the 95th percentile of the square of the standard normal distribution.

2. the 95th percentile of the chi-squared distribution is equal to the 97.5th percentile of the square of the standard normal distribution.

3. the 97.5th percentile of the chi-squared distribution is equal to the 98.75th percentile of the square of the standard normal distribution.

```{r}
```
Q2.2
(b)
```{r}
v1 <- 1
v2 <- 20

# 95th percentile of square of t distribution
qt(0.95, df=v2)^2

# 90th percentile of f distribution (1, v2)
qf(0.90, df1=v1, df2=v2)

# 97.5th percentile of square of t distribution
qt(0.975, df=v2)^2

# 95th percentile of f distribution (1, v2)
qf(0.95, df1=v1, df2=v2)

# 98.75th percentile of square of t distribution
qt(0.9875, df=v2)^2

# 97.5th percentile of f distribution (1, v2)
qf(0.975, df1=v1, df2=v2)
```
We can see from the following results that:

1. the 90th percentile of the F distribution (1, v2) is equal to the 95th percentile of the square of the t distribution (v2).

2. the 95th percentile of the F distribution (1, v2) is equal to the 97.5th percentile of the square of the t distribution (v2).

3. the 97.5th percentile of the F distribution (1, v2) is equal to the 98.75th percentile of the square of the t distribution (v2).
```{r}
```
Q2.4
(a)
```{r}
weekof <- c("January 30", "June 39", "March 2", "October 26", "February 7")
no_of_cars_sold_y <- c(20, 18, 10, 6, 11)
ave_no_of_salesppl_x <- c(6, 6, 4, 2, 3)

# create matrix with the data
data <- cbind(weekof, no_of_cars_sold_y, ave_no_of_salesppl_x)

# create data frame
df <- as.data.frame(data)
transform(df, no_of_cars_sold_y = as.numeric(no_of_cars_sold_y), ave_no_of_salesppl_x = as.numeric(ave_no_of_salesppl_x))

# construct scatterplot of y against x, where y is the number of cars sold and x is the average number of salespeople
plot(ave_no_of_salesppl_x, no_of_cars_sold_y, xlab="Ave # of salespeople", ylab="# of cars sold", main="Scatterplot of # of cars sold against ave # of salespeople")
```
Q2.4
(b)
```{r}
# estimate intercept using Method of least squares
xbar = mean(ave_no_of_salesppl_x)
ybar = mean(no_of_cars_sold_y)
b1 = sum((ave_no_of_salesppl_x - xbar) * (no_of_cars_sold_y - ybar)) / sum((ave_no_of_salesppl_x - xbar)^2)
b0 = ybar - b1 * xbar
cat("The estimated intercept:", b0, "\n")
cat("The estimated slope of the least squares line:", b1, "\n")

# check using built in function 'lm'
model <- lm(no_of_cars_sold_y ~ ave_no_of_salesppl_x)

# print the summary of the model
summary(model)
```
Q2.4
(c)
```{r}
# plot the least squares line
plot(ave_no_of_salesppl_x, no_of_cars_sold_y, xlab="Ave # of salespeople", ylab="# of cars sold", main="Scatterplot of # of cars sold against ave # of salespeople")
abline(model, col="red")
```
Q2.4
(d)
```{r}
# estimate the number of cars sold when the average number of salespeople is 5
x = 5
y = b0 + b1 * x
cat("The estimated number of cars sold when the average number of salespeople is 5:", y, "\n")
```
Q2.4
(e)
```{r}
# Calculate the fitted values mu_hat for each observed value of x
mu_hat = b0 + b1 * ave_no_of_salesppl_x

# Calculate the residuals e_i for each observed value of x
e_i = no_of_cars_sold_y - mu_hat

# Plot the residuals against the fitted values
plot(mu_hat, e_i, xlab="Fitted values", ylab="Residuals", main="Residuals against Fitted values")

```
Q2.4
(f)
```{r}
# Calculate manually an estimate of the variance sigma square
n = length(no_of_cars_sold_y)
rss = sum(e_i^2)
estimate_variance = rss / (n - 2)
cat("The estimated variance:", estimate_variance, "\n")
```
```{r}
```
Q2.6
(a)
```{r}
# import data
q6data = read.table(file='hooker.txt', header = T)
q6data
```
```{r}
# Plot TEMP against AP, where BT is the temperature and AP is the air pressure
plot(q6data$AP, q6data$BT, xlab="AP", ylab="TEMP", main="Scatterplot of TEMP against AP")
```
- A Linear model seems appropriate as the relationship between TEMP and AP seems to be linear.

Q2.6
(b)
```{r}
x = 100 * log(q6data$AP, base = exp(1))

# plot the scatterplot of TEMP against x
plot(x, q6data$BT, xlab="100 * ln(AP)", ylab="TEMP", main="Scatterplot of TEMP against 100 * ln(AP)")
```
Q2.6
(c)
```{r}
xbar = mean(q6data$AP)
ybar = mean(q6data$BT)
b1 = sum((q6data$AP - xbar) * (q6data$BT - ybar)) / sum((q6data$AP - xbar)^2)
b0 = ybar - b1 * xbar
cat("The estimated intercept:", b0, "\n")
cat("The estimated slope:", b1, "\n")

# check using built in function 'lm'
q6model <- lm(q6data$BT ~ q6data$AP)

# print the summary of the model
summary(q6model)

```
```{r}
# plot the least squares line
plot(q6data$AP, q6data$BT, xlab="AP", ylab="TEMP", main="Scatterplot of TEMP against AP")
abline(q6model, col="red")
```