---
Title: Assignment 2
Author: Zane Yee Sun 
ID: 124400004
---
# Assignment 2 - Linear Models
### Q2.6 
```{r}
# set your path
data<-read.table(file = "hooker.txt",header = TRUE)
head(data)
```
```{r}
# Initialize variables from HW1 data
TEMP<-data$BT
AP<-data$AP

x<-100*log(AP)

x_mean<-mean(x)
y_mean<-mean(TEMP)

Sxx<-sum((x-x_mean)^2)
Sxy<-sum((x-x_mean)*TEMP)

beta_1<-Sxy/Sxx
beta_0<- y_mean-beta_1*x_mean
```
#### (d)(i)
```{r}
# 95% confidence interval for beta_1
n<-length(x)
alpha<-0.05
t<-qt(1-alpha/2,df=n-2)
SE<-sqrt(sum((TEMP-beta_0-beta_1*x)^2)/(n-2))/sqrt(Sxx)
CI<-c(beta_1-t*SE,beta_1+t*SE)
cat("95% confidence interval for beta_1 is: ",CI)
```
#### (d)(ii)
```{r}
# Ave temperature when AP = 25
x_25<-100*log(25)
y_25<-beta_0+beta_1*x_25
cat("Average temperature when AP = 25 is: ",y_25)
```
### Q2.8
```{r}
# Initialize data from question
company <- c("General Motors", "Ford/Volvo", "Renault/Nissan", "Volkswagen", "DaimlerChrysler", "Toyota", "Fiat", "Honda", "PSA", "BMW")
cars_sold <- c(8149, 7316, 4778, 4580, 4506, 4454, 2535, 2291, 2278, 1187)
revenue <- c(1996, 2118, 1174, 943, 1813, 1175, 628, 605, 465, 447)

# Create data frame
df <- data.frame(company, cars_sold, revenue)

# Fit linear model y = revenue, x = cars sold (sales)
fit <- lm(revenue ~ cars_sold, data = df)

# Print summary
summary(fit)
```
#### (a)
Hypothesis testing for the importance of cars sold in predicting revenue.
- Null hypothesis: The slope of the linear model is 0.
- Alternative hypothesis: The slope of the linear model is not 0.

Since the p-value = 0.000156 < 0.05, we reject the null hypothesis that the slope is 0. There is a significant linear relationship between revenue and cars sold.
---
#### (b)
```{r}
# 95% confidence interval for the regression coefficient of the number of cars sold
n <- length(cars_sold)
alpha <- 0.05
t <- qt(1-alpha/2, df = n-2)
beta_0 <- coef(fit)[1]
beta_1 <- coef(fit)[2]
x_mean <- mean(cars_sold)
y_mean <- mean(revenue)
Sxx <- sum((cars_sold - x_mean)^2)
SE <- sqrt(sum((revenue - beta_0 - beta_1 * cars_sold)^2) / (n-2)) / sqrt(Sxx)
CI <- c(beta_1 - t * SE, beta_1 + t * SE)
cat("95% confidence interval for beta_1 is: ", CI)
```
```{r}
# Checking the 95% confidence interval using confint()
confint(fit, level = 0.95)
```
#### (c)
```{r}
# 90% confidence interval for the regression coefficient of the numbers of cars sold
alpha <- 0.1
t <- qt(1-alpha/2, df = n-2)
SE <- sqrt(sum((revenue - beta_0 - beta_1 * cars_sold)^2) / (n-2)) / sqrt(Sxx)
CI <- c(beta_1 - t * SE, beta_1 + t * SE)
cat("90% confidence interval for beta_1 is: ", CI)
```
```{r}
# Check again using confint()
confint(fit, level = 0.90)
```
#### (d)
```{r}
# get the coefficent of determination
R2 <- summary(fit)$r.squared
cat("The coefficient of determination is: ", R2)
```
#### (e)
```{r}
# Calculate SD of revenue after factoring the sales of cars
y_hat <- beta_0 + beta_1 * cars_sold
SSE_withx <- sum((revenue - y_hat)^2)

# Calculate SD of revenue without factoring the sales of cars
y_mean <- mean(revenue)
SSE_withoutx <- sum((revenue - y_mean)^2)

# Output and compare the two SSEs
cat("SSE with x: ", SSE_withx, "\n")
cat("SSE without factoring x: ", SSE_withoutx, "\n")
```
#### (f)
```{r}
# Calculate estimates for BMW
cars_sold_BMW <- 1187
revenue_BMW <- beta_0 + beta_1 * cars_sold_BMW
cat("The estimated revenue for BMW is: ", revenue_BMW)
```
### Q2.14

#### (a)
Let 'x' denote s the quantities of calcium in carefully prepared solutions.
Let 'y' denote the corresponding analytical results.
```{r}
# Initialize x and y variables
x <- c(4, 8, 12.5, 16, 20, 25, 31, 36, 40, 40)
y <- c(3.7, 7.8, 12.1, 15.6, 19.8, 24.5, 31.1, 35.5, 39.4, 39.5)

# Manually fit the linear model of y and x
x_mean <- mean(x)
y_mean <- mean(y)
Sxx <- sum((x - x_mean)^2)
Sxy <- sum((x - x_mean) * y)
beta_1 <- Sxy / Sxx
beta_0 <- y_mean - beta_1 * x_mean

# Print the coefficients
cat("The estimated coefficients are: beta_0 = ", beta_0, ", beta_1 = ", beta_1, "\n")
```
```{r}
# Fit a linear model using lm()
fit <- lm(y ~ x)

summary(fit)
```
The assumptions made:
1. The data is normally distributed.
2. Each instance of x_i is independent of other instances, and the same goes for y_i.
---
#### (b)



